---
title: "Lecture 8 (VIZ + EDA II)"
author: "CJ Snyder"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Note: "theme_set" is a general code to apply the two agruments to whatever ggplot you want

# Loading Dataset from NOAA
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

# Creating Initial Plot
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color=name), alpha=.5)
```

# Adding Labels to Above _ggplot_
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color=name), alpha=.5) +
  labs(
    title="Temperature Plot",
    x="Minimum Daily Temperature (C)",
    y="Maximum Daily Temperature (C)",
    caption="Data from NOAA Packa") + 
  scale_x_continuous(
    breaks=c(-15, 0, 15),
    labels=c("-15 C", "0", "15"))
```

# Transformation of Plot
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color=name), alpha=.5) +
  labs(
    title="Temperature Plot",
    x="Minimum Daily Temperature (C)",
    y="Maximum Daily Temperature (C)",
    caption="Data from NOAA Packa") + 
  scale_x_continuous(
    breaks=c(-15, 0, 15),
    labels=c("-15 C", "0", "15")) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

# Colors and Themes in _ggplot_
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color=name), alpha=.5) +
  labs(
    title="Temperature Plot",
    x="Minimum Daily Temperature (C)",
    y="Maximum Daily Temperature (C)",
    caption="Data from NOAA Package") +
  scale_color_hue(name="Location", h=c(100,300))
```

## Changing color sheme using _viridis_ package
```{r}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_temp_plot
```

# Changing Themes
```{r}
ggp_temp_plot +
  theme_classic() +
  theme(legend.position = "bottom")

ggp_temp_plot +
  theme_bw() +
  theme(legend.position = "bottom")

ggp_temp_plot +
  theme_minimal() +
  theme(legend.position = "bottom")

ggp_temp_plot +
  ggthemes::theme_excel() +
  theme(legend.position = "bottom")
```

Note: Overall themes should be coded/ordered first, then detail-specific (i.e. having theme_bw come before theme(...))

# Including data agruments in _geom_point/line__ggplot_
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

## Brief sidenote about colors
```{r}
waikiki %>% 
  ggplot(aes(x=date, y=tmax))+
  geom_point(color="green")
```
Note: need to include color statment in point geometry


# Multi-panel Plots via _patchwork_
```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p

tmax_tmin_p + prcp_dens_p + tmax_date_p

tmax_tmin_p / (prcp_dens_p + tmax_date_p)
```


# Data Manipulation in _ggplot_
```{r}
# Original ggplot
weather_df %>% 
  ggplot(aes(x=name, y=tmax, color=name))+
  geom_boxplot()

# Changing the order of x variable
weather_df %>% 
  mutate(
    name=factor(name),
    name=fct_relevel(name,"Waikiki_HA", "CentralPark_NY")) %>% 
  ggplot(aes(x=name, y=tmax, color=name))+
  geom_boxplot()

# Reordering instead of releveling
weather_df %>% 
  mutate(
    name=factor(name),
    name=fct_reorder(name, tmax)) %>% 
  ggplot(aes(x=name, y=tmax, color=name))+
  geom_boxplot()
```


# Random graph
```{r}
weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```



